### 2020-01-18 code/02-Master.Rmd
Preamble Work

---
title: "AD Target Gene Report"
author: "Jake Gockley"
date: "11/11/2019"
header-includes:
  - \usepackage{multicol}
  - \newcommand{\btwocol}{\begin{multicols}{2}}
  - \newcommand{\etwocol}{\end{multicols}}
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(readr)
library(biomaRt)
library(knitr)
library(kableExtra)
library(tidyverse)
library(grid)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(yaml)
```

```{r yamlIMPT, eval = T, include=T}
#Assign your YAML formated config file to a varable
configuration <- "YAML_FILE"
config <- read_yaml(configuration)

setwd(config$filedir)
plotdir=paste0( 'runs/', config$runname, '/figures' )
tabledir=paste0( 'runs/', config$runname, '/tables' )

Tissues <- config$tissue

#DE  = 'syn14237651'
#eQTL_Cer = 'syn16984411'
#eQTL_TCX = 'syn16984410'
#eQTL_DLPFC = 'syn16984409'
#eQTL_EasyYesNo = 'syn12514912
#AgoraDrugability = 'syn13363443'
#Proteomics = 'syn18689335'
#AD Associated Mets Networks: syn11932957
#AD Meta Network Modual Associations syn11954640
```
```{r Module1, eval = T, include=T}

```