### 2022-02-25 code/RNA-Seq_Meta_Analysis.Rmd
update the weights

---
title: "Meta analysis of reprocessed RNASeq data from AMP-AD for hg38 Reprocessed Data (all 7 brain regions)"
author: "Adapted from Thanneer Perumal's code"
date: "`r date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
## Load required libraries
library(CovariateAnalysis)
library(data.table)
library(plyr)
library(tidyverse)
library(knitr)
library(githubr)
library(doParallel)
library(foreach)

options(xtable.type="html")
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
synapser::synLogin()
parentId = 'syn22351719';
activityName = 'Meta analysis of differential expression hg38 Reprocessed RNASeq';
activityDescription = 'Fixed and random effect meta-analysis of AMP-AD data hg38 Reprocessed Data (6 brain regions - No CBE)';
thisFileName <- 'RNA-Seq_Meta_Analysis.Rmd'
# Github link
thisRepo <- getRepo(repository = "Sage-Bionetworks/AD_TargetRank", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

## Data download
### Get study specific covariates and adjusted expression from synapse

```{r get.exp}
# Function to read tsv/csv files from synapse
downloadFile <- function(id){ 
  fread(synapser::synGet(id)$path, data.table = F)
}
# Read covariates
#covar.ids = c(ROSMAP_DLPFC = 'syn26340637 ', 
#              ROSMAP_PCC = 'syn26350606',
#              ROSMAP_ACC = 'syn26374583 ',
#              MSSM_FP = 'syn26453689', 
#              MSSM_IFG = 'syn26454154', 
#              MSSM_PHG = 'syn26453943', 
#              MSSM_STG = 'syn26454639', 
#              MAYO_TCX = 'syn26374583')

covar.ids = c(ROSMAP = 'syn26967450 ', 
              MSSM = 'syn27068753', 
              MAYO = 'syn27024950')
#_tissue

all.used.ids = as.character(covar.ids)
covar =  covar.ids %>%
  lapply(downloadFile)
#covar$ROSMAP_DLPFC$tissue <- 'DLPFC'
#covar$ROSMAP_PCC$tissue <- 'PCC'
#covar$ROSMAP_ACC$tissue <- 'ACC'
#covar$MSSM_FP$tissue <- 'FP'
#covar$MSSM_IFG$tissue <- 'IFG'
#covar$MSSM_PHG$tissue <- 'PHG'
#covar$MSSM_STG$tissue <- 'STG'
#covar$MAYO_TCX$tissue <- 'TCX'

covar = covar %>%
  data.table::rbindlist(use.names = T, fill = T, idcol = 'Study')

# Read adjusted expression
#expr.ids = c(ROSMAP_DLPFC = 'syn26340641', 
#              ROSMAP_PCC = 'syn26374587',
#              ROSMAP_ACC = 'syn26350610',
#              MSSM_FP = 'syn26453693', 
#              MSSM_IFG = 'syn26454158', 
#              MSSM_PHG = 'syn26453947', 
#              MSSM_STG = 'syn26454643', 
#              MAYO_TCX = 'syn26387667')
expr.ids = c(ROSMAP_DLPFC = 'syn26967454', 
              MSSM_STG = 'syn27068759', 
              MAYO_TCX = 'syn27024966')

all.used.ids = c(all.used.ids, as.character(expr.ids))

expr =  expr.ids %>%
  lapply(downloadFile) %>%
  plyr::join_all(type = 'full')
#expr <- as.data.frame(expr) %>%
#  column_to_rownames('feature')

```

## Perform fixed and random effect meta-analysis for each gene across 7 different brain regions
#### Meta-analysis of AD-CONTROL
```{r ad.cntrl}
# Mean and sd of gene expression in each tissue x diagnosis group
tissue.dx.summary = plyr::ddply(covar, .(tissue, diagnosis), .fun = function(x, y){
  data.frame(ensembl_gene_id = y$feature,
             n = dim(x)[1],
             mn = rowMeans(y[,x$specimenID], na.rm = T),
             sd = apply(y[,x$specimenID], 1, sd, na.rm = T))
}, expr)
# Perform meta-analysis for AD-CONTROL comparison - Exclude CBE
meta.anlz.ad_cntrl = plyr::ddply(tissue.dx.summary[(tissue.dx.summary$tissue%in%'CBE')==F,], .(ensembl_gene_id), .fun = function(x){
  exp.effect = dplyr::filter(x, diagnosis == 'AD')
  rownames(exp.effect) = exp.effect$tissue
  cntrl.effect = dplyr::filter(x, diagnosis == 'CT')
  rownames(cntrl.effect) = cntrl.effect$tissue
  cntrl.effect = cntrl.effect[rownames(exp.effect), ]

  tmp = metacont(exp.effect$n, exp.effect$mn, exp.effect$sd, 
                 cntrl.effect$n, cntrl.effect$mn, cntrl.effect$sd,
                 studlab = exp.effect$Tissue,
                 sm = 'SMD', method.smd = 'Hedges',
                 method.tau = 'REML')

  return(data.frame(tmp[c('TE.fixed', 'seTE.fixed', 'lower.fixed', 'upper.fixed', 'zval.fixed', 'pval.fixed',
                   'TE.random', 'seTE.random', 'lower.random', 'upper.random', 'zval.random', 'pval.random',
                   'Q', 'tau', 'H', 'I2')]))
}, .parallel = TRUE, .paropts = list(.packages = c('meta', 'dplyr'))) %>%
  dplyr::mutate(fdr.fixed = p.adjust(pval.fixed, method = 'fdr'),
                fdr.random = p.adjust(pval.random, method = 'fdr'))
p = list()
p[[1]] = ggplot(meta.anlz.ad_cntrl, aes(x = -log10(fdr.fixed), y = -log10(fdr.random)))+geom_point()
p[[2]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.fixed), x = TE.fixed))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
p[[3]] = ggplot(meta.anlz.ad_cntrl, aes(y = -log10(fdr.random), x = TE.random))+geom_point()+geom_hline(yintercept = -log10(0.05), color = 'red')
ggpubr::ggarrange(plotlist = p, ncol = 3, nrow = 1)
# Get gene sets
tmp = meta.anlz.ad_cntrl %>%
  dplyr::select(ensembl_gene_id, TE.fixed, TE.random, fdr.fixed, fdr.random) %>%
  tidyr::gather(Feature, Value, -ensembl_gene_id) %>%
  tidyr::separate(Feature, c('Feature1', 'Model'), sep = '\\.') %>%
  tidyr::spread(Feature1, Value) %>%
  dplyr::filter(fdr <= 0.05, abs(TE) >= 0.2)
tmp$Direction[tmp$TE <= 0] = 'DOWN'
tmp$Direction[tmp$TE >= 0] = 'UP'
gs = tmp %>%
  plyr::dlply(.(Model, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })
names(gs) = paste0('ad.control.', names(gs))
all.gs = gs
```

#### Adjust the AD-Control Model for weights
```{r weights}
meta.anlz.ad_cntrl <- meta.anlz.ad_cntrl[complete.cases(meta.anlz.ad_cntrl),]
meta.anlz.ad_cntrl$TE.random.abs <- abs( meta.anlz.ad_cntrl$TE.random )

#Assign Sig Or No
meta.anlz.ad_cntrl$Sig <- 'NO'
meta.anlz.ad_cntrl[meta.anlz.ad_cntrl$fdr.random < 0.05, ]$Sig <- 'YES'

meta.anlz.ad_cntrl <- meta.anlz.ad_cntrl[ order(meta.anlz.ad_cntrl$TE.random.abs), ]
meta.anlz.ad_cntrl$y <- c( 1:dim(meta.anlz.ad_cntrl)[1] )/dim(meta.anlz.ad_cntrl)[1]

meta.anlz.ad_cntrl$Type <- 'Actual Rank'

mylogit <- glm(meta.anlz.ad_cntrl$y  ~ abs( meta.anlz.ad_cntrl$TE.random.abs), data = meta.anlz.ad_cntrl, family = "binomial")

#Make Weight Template
Work <- meta.anlz.ad_cntrl[, c('ensembl_gene_id','TE.random', 'seTE.random','fdr.random','TE.random.abs')]
Work$log_abs_PVal <- abs( meta.anlz.ad_cntrl$TE.random.abs )
colnames(Work)[ colnames(Work) == 'ensembl_gene_id' ] <- 'ENSG'

#Calculate predicted Weights
Work2 <- Work  
Work2$Type <- "Predicted Weight"
#Work2 <- Work2[ , c('ENSG', 'log_abs_PVal',"Type") ]
Work2$y <- 1/(1+exp( -( mylogit$coefficients[1]+mylogit$coefficients[2]* abs( meta.anlz.ad_cntrl$TE.random.abs ) ) ))
Work2$PreAdjWeight <- Work2$y

#Add Actual Ranks to Template
Work$y <- c( 1:dim(Work)[1] )/dim(Work)[1]
Work$Type <- 'Actual Rank'
mylogit <- glm(Work$y  ~ abs( Work$TE.random.abs), data = Work, family = "binomial")
Work$PreAdjWeight <- Work$y

Weights <- as.data.frame( rbind(Work,Work2) )
Weights[ Weights$fdr.random > 0.05,]$y <- 0
Weights <- Weights[ order(-Weights$y), ]

G_list <- fread(synapser::synGet('syn26453691')$path)
all.used.ids <- c(all.used.ids,'syn26453691')

G_list <- as.data.frame(G_list)
G_list$ensembl_gene_id <- as.character(G_list$ensembl_gene_id)
G_list$hgnc_symbol <- as.character(G_list$hgnc_symbol)

G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]

G_list <- G_list[ !duplicated(G_list$ensembl_gene_id),]

row.names(G_list) <- G_list$ensembl_gene_id
Weights$Gene <- G_list[ as.character(Weights$ENSG),]$hgnc_symbol

Weights <- Weights[, c( "ENSG", "Gene", "TE.random", "seTE.random", "fdr.random", 'TE.random.abs', "log_abs_PVal", "Type", "PreAdjWeight", "y" )]

Weights$plotname <- NA
label_name <- c( "MTCH1", "MTOR", "APPL1", "BAX", "STMN2", "SLC9A9", "TOMM40L", "GABRA2",  "MTCH2"  )
Weights[ Weights$Gene %in% label_name, ]$plotname <- Weights[ Weights$Gene %in% label_name, ]$Gene

A <- Weights[ Weights$Type == 'Actual Rank', ]
row.names(A) <- A$ENSG
colnames(A)[colnames(A)=='PreAdjWeight'] <- 'PreAdjWeight_Actual'
colnames(A)[colnames(A)=='y'] <- 'y_Actual'

B <- Weights[ Weights$Type == 'Predicted Weight', ]
row.names(B) <- B$ENSG
colnames(B)[colnames(B)=='PreAdjWeight'] <- 'PreAdjWeight_Predicted'
colnames(B)[colnames(B)=='y'] <- 'y_Predicted'

plotframe <- as.data.frame(cbind(A[,c('ENSG','Gene', 'TE.random', 'seTE.random', 'fdr.random', 'TE.random.abs', 'log_abs_PVal', 'PreAdjWeight_Actual','y_Actual', 'plotname')], B[row.names(A), c('PreAdjWeight_Predicted','y_Predicted')]))

plotframe <- plotframe[ plotframe$fdr.random < 0.05, ]
plotframe$logTE <- log(abs(plotframe$TE.random.abs))

library(ggplot2)
library(ggrepel)
ggplot2::ggplot(data=plotframe) + 
  ggplot2::geom_point( ggplot2::aes(x=logTE, y=y_Predicted)) +
  ggrepel::geom_label_repel(data = plotframe, 
                            ggplot2::aes(x=logTE, y=y_Predicted, label = plotname, size = NULL, color = NULL),
                            nudge_x = .45,
                            nudge_y = -.05,
                            segment.size  = 0.2,
                            segment.color = "grey50",
                            direction     = "x"
) +
  ggplot2::xlab("Log2( absolute(Effect Size))") + 
  ggplot2::ylab("Assigned weight") +
  ggplot2::labs(title="Weights of Differentially Expressed Transcripts") + 
  ggplot2::theme(plot.title =  ggplot2::element_text(hjust = 0.5))

```

## Store results in synapse
```{r syn.store, cache = FALSE, include=FALSE}
CODE <- synapser::synStore(synapser::Folder(name = "MetaAnalysis", parentId = parentId))
Syns_Used <- all.used.ids

# Write results to files
fwrite(meta.anlz.ad_cntrl, file = 'RNASeq_meta.anlz.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta.anlz.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis across 6 brain regions (No CBE) hg38 Reprocessed Data', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)

# Write results to files
fwrite(Weights, file = 'RNASeq_meta_weights.ad_cntrl.tsv', sep = '\t', row.names = F, quote = F)
ENRICH_OBJ <-  synapser::synStore( synapser::File( path='RNASeq_meta_weights.ad_cntrl.tsv', name = 'RNASeq AD-Control meta-analysis Weights', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
```

## Source code